---
title: "PolNet Intro Network Analysis 2022"
author: "Alex Genovese"
date: '2022-06-13'
output: html_document
---

# Terms

* Network: A collection of **nodes** joined by **edges**
  
* Node: Unit we are examining
  + Could be countries, people, cities, etc.
  + Synonyms: vertex, actors
  
* Edge: Connection between two nodes

* One-mode networks: all actors are tied to one another according to one relation (e.g. friendship)

* Two-mode networks: Actors are tied to particular events

* Types of Networks:
  + Undirected, binary
  + Directed, binary
  + Undirected, Valued
  + Directed, Valued
  
* Data: Adjenecy matrix vs. edge list
  + Adjency matrix - values show connection between nodes (often omit 0, just care about 1 if binary)
  + Edge list - 2 columns, node 1/node 2 (Sometimes 3rd column for weight if not binary; usually missing edges not included)
  
  

# iGraph R
```{r iGraph Setup, include=FALSE}
library(igraph)
library(rio)
```
```{r iGraph, include=FALSE}
# ?graph_from_data_frame
node <- import("nodeList.csv")
edge <- import("edgeList.csv")

View(node)
View(edge)

## Directed, unweighted network plot
net_igraph <- graph_from_data_frame(d=edge, v=node, directed=T)

net_igraph_eonly <- graph_from_data_frame(d=edge, directed=T)

net_igraph 
net_igraph_eonly # No isolates: they are the same

plot(net_igraph,
     edge.arrow.size=.5,
     edge.curved=0.2,
     edge.color="black",
     edge.width=1.5,
     vertex.size=5, 
     vertex.frame.color="gray",
     vertex.color="white",
     vertex.label.cex=1,
     vertex.label.dist=1.2,
     vertex.label.color="black")


## Undirected, unweighted
net_igraph2<-graph_from_data_frame(d=edge, v=node, directed=F)
net_igraph2 # duplicates

plot(net_igraph2,
     edge.color="black",
     edge.width=1.5,
     vertex.size=5, 
     vertex.frame.color="gray",
     vertex.color="white",
     vertex.label.cex=1,
     vertex.label.dist=1.2,
     vertex.label.color="black")

net_igraph3 <- as.undirected(net_igraph, mode='collapse')
net_igraph3

plot(net_igraph3,
     edge.color="black",
     edge.width=1.5,
     vertex.size=5, 
     vertex.frame.color="gray",
     vertex.color="white",
     vertex.label.cex=1,
     vertex.label.dist=1,
     vertex.label.color="black")

## Directed, Weighted 
net_igraph
E(net_igraph)$Weight

### Add vertex/edge attributes
V(net_igraph)$gender <- c("M","F","F","M","M","M")
V(net_igraph)$gender

E(net_igraph)
E(net_igraph)$color <- ifelse(E(net_igraph)$Weight==1, "darkred", "navy") 

plot(net_igraph,
     edge.width=1.5*E(net_igraph)$Weight,
     edge.arrow.size=.6,
     edge.curved=0.2,
     edge.color=E(net_igraph)$color,
     vertex.size=7, 
     vertex.frame.color="white", 
     vertex.color=ifelse(V(net_igraph)$gender=="M", "gray", "black"),
     vertex.label.cex=1.5,
     vertex.label.dist=1.5)

## Undirected, Weighted 
net_igraph2 ## not collapsed
E(net_igraph2)$Weight

### Add vertex/edge attributes
V(net_igraph2)$gender <- c("M","F","F","M","M","M")
V(net_igraph2)$gender

E(net_igraph2)
E(net_igraph2)$color <- ifelse(E(net_igraph2)$Weight==1, "darkred", "navy") 

plot(net_igraph2,
     edge.width=1.5*E(net_igraph)$Weight,
     #edge.arrow.size=.6,
     #edge.curved=0.2,
     edge.color=E(net_igraph2)$color,
     vertex.size=8, 
     vertex.frame.color="white", 
     vertex.color=ifelse(V(net_igraph2)$gender=="M", "gray", "black"),
     vertex.label.cex=1.5,
     vertex.label.dist=1.5)
```
```{r iGraph Example, include=FALSE}
## Real data: defense pacts (ATOP: Leeds et al. 2002)
atop <- import("atop_sample2.csv")

head(atop)
atop1997_dat <- subset(atop, year==1997, c(stateabb1, stateabb2))

atop1997 <- graph.data.frame(atop1997_dat)
atop1997und <- as.undirected(atop1997, mode='collapse')  

# Undirected, unweighted

plot(atop1997und,
     vertex.label=NA,
     edge.width=1.5, 
     vertex.size=2.5, 
     vertex.color="blue",
     layout=layout_with_kk)

plot(atop1997und,
     #vertex.label=NA,
     vertex.label.cex=0.6, 
     vertex.label.color="black",
     vertex.label.dist=1,
     edge.width=1.5, 
     vertex.size=2.5, 
     vertex.frame.color="black", 
     vertex.color="blue",
     layout=layout_with_kk)

```
# Statnet R
```{r Statnet Setup, include=FALSE}
detach("package:igraph", unload = TRUE)
library(statnet)
```
```{r Statnet, include=FALSE}
head(edge)

# ?network
net_statnet <- network(edge, matrix.type="edgelist") 
#matrix.type="adjacency"
#directed=T

net_statnet
class(net_statnet)

plot(net_statnet, 
     displaylabels=T)

## Undirected
# net_statnet2 <- network(edge, matrix.type = "edgelist", directed = FALSE) ## Error due to duplicates 

net_statnet3 <- network(edge, 
                        matrix.type = "edgelist",
                        directed = FALSE,
                        multiple = TRUE)

net_statnet3

## Weights

net_statnet%e%"Weight"

plot(net_statnet,
     displaylabels=T,
     edge.lwd=5*net_statnet%e%"Weight") 

netweighted <- network(edge, 
                       matrix.type="edgelist",
                       ignore.eval=FALSE) ## Another way to load weights

netweighted
#ignore.eval: logical; ignore edge values?


plot(netweighted,
     displaylabels=T,
     edge.lwd=5*netweighted%e%"Weight")

netweighted[,] #adjacency matrix without weight
as.sociomatrix.sna(netweighted,"Weight")
#as.sociomatrix: Coerce One or More Networks to Sociomatrix Form

### Add vertex/edge attributes
netweighted %v% "gender" <- c("M","F","F","M","M","M")
netweighted %v% "gender"

netweighted %v% "vertex.names"

netweighted %e% "color" <- ifelse(netweighted %e% "Weight"==1, "darkred",
                                  "navy")
netweighted %e% "color"

plot(netweighted,
     displaylabels=T,
     edge.lwd=5*netweighted%e%"Weight",
     edge.col=netweighted%e%"color",
     vertex.col=ifelse(netweighted%v%"gender"=="F","black","gray"))


```
# Network statistics

## Network Density
* Density refers to the proportion of ties in a network that are actually present
  + In undirected networks: $d =\frac{L}{n(n-1)/2}$ where L refers to the actual number of lines present in the network and n to the number of nodes present in the network 
  + In directed networks: $d = \frac{L}{n(n-1)}$

## Subgraphs
* Dyad
  + Is the relationship reciprocal?
  + Mutual, asymmetric, or null?
* Triad
  + Transitivity
  + A to B, B to C, and A to C
  + "States are more likely to send diplomatic missions to countries where their diplomatic partners also send missions" (Kinne 2014)

## Research Design
* Network as IV
  + Friendship -> Behavior
  + Centrality -> Promotion

* Network as DV
  + Homophily -> Friendship
  + Personality -> Central actors

## Centrality
* Why do we care about central nodes?
  + Relational ties (linkages) between actors are channels for transfer or "flow" of resources.
  + Network models focusing on individuals view the network structural environment as providing opportunities for or constraints on individual action (Wasserman and Faust 1994)
* In terms of information flows?
  + Get the information earlier than others
  + Control the flow of information
* What does it mean for an actor to be central?
  + Prominent, influential
  + Having great visibility, prestige, power, control, autonomy
  + Leaders, gatekeepers
  + "For non-negative relations such as friendship or trust, centrality tends to be viewed as a positive thing for nodes, providing actors with opportunities to influence others and receive flows (i.e., information)."
* Degree centrality
  + Total number of ties: $d_{i} = \Sigma_{j}x_{ij}$
* Eigenvector centrality
  + Proportional to the sum of centralities of the nodes it is connected to
  + If one's alters have high degree centrality, then the focal actor has high eigenvector centrality
  + $e_{i} = \lambda\Sigma_{j}x_{ij}e_{j}$
  + Ego: focal actor
  + Alters: the nodes that ego is connected
* Betweenness Centrality
  + How often an actor rests between two other actors
  + $C_{B}(k) = \Sigma\frac{\alpha_{ikj}}{\alpha_{ij}}, i\neq j \neq k,$ where $\alpha_{ikj}$ is the number of geodesics linking actors i and j that pass through node k and $\alpha_{ij}$ is the number of geodesics linking actors k and j (geodesic means the shortest path in the network)
* Closeness Centrality
  + Emphasizes an actor's independence
  + If an actor is not central, the actor generally needs to rely on others to relay messages through the network. Thus, an actor who is close to many other actors is a very independent actor.



# Network statistics and centrality R
```{r Network Statistics and Centrality Statnet, eval=FALSE, include=FALSE}
# Calculate Network Statistics

## Using statnet (count edges & triangles)
summary(net_statnet ~ edges)
atopnet<-network(atop1997_dat, matrix.type="edgelist")
summary(atopnet ~ edges+triangles)
```
```{r Network Statistics and Centrality iGraph}
## Using igraph
detach("package:statnet", unload = TRUE)
library(igraph)

net_igraph<-graph_from_data_frame(d=edge, v=node, directed=T)

## Dyad
plot(net_igraph)
summary(net_igraph)

# ?dyad.census
#mut: The number of pairs with mutual connections.
#asym: The number of pairs with non-mutual connections.
#null: The number of pairs with no connection between them.

dyad.census(net_igraph)

## Triads
# ?triad.census
triad.census(net_igraph)
### E.g.030T: a->e<-b, a->b
###           a->d<-b, a->b

## Undirected
net_igraph3 <- as.undirected(net_igraph, mode='collapse')

dyad.census(net_igraph3) # number of ties
triad.census(net_igraph3) 

## Comparing density
atop1997 <- graph.data.frame(atop1997_dat)
atop1997und <- as.undirected(atop1997, mode='collapse')

atop2002_dat <- subset(atop, year==2002, c(stateabb1, stateabb2))
atop2002 <- graph.data.frame(atop2002_dat)
atop2002und <- as.undirected(atop2002, mode='collapse')

atop1997und 
atop2002und 

edge_density(atop1997und)
edge_density(atop2002und)

# Centrality
## Degree - Number of adjacent ties for a node
?degree

degree(net_igraph,
       mode = "in")

degree(net_igraph,
       mode = "out")

degree(net_igraph,
       mode = "total")

### ATOP

plot(atop1997und,
     #vertex.label=NA,
     vertex.label.cex=0.6, 
     vertex.label.color="black",
     vertex.label.dist=1,
     edge.width=1.5, 
     vertex.size=2.5, 
     vertex.frame.color="black", 
     vertex.color="blue",
     layout=layout_with_kk)

degree(atop1997und)

### Eigenvector (undirected)
eigen_centrality(atop1997und)$vector

### Betweenness (directed or undirected)
betweenness(net_igraph)
betweenness(atop1997und, directed = FALSE)


### Closeness (directed or undirected)
closeness(net_igraph)
closeness(atop1997und)



```








